---
title: 'Social Network Analytics Empirical Exercise #1'
output:
  html_document: default
  pdf_document: default
---
Jent LaPalm

```{r, warning=FALSE, message=FALSE}
library(igraph)
library(data.table)
library(randtests)
library(stringr)
library(Matrix)
library(expm)

path <- "https://raw.githubusercontent.com/jentlapalm/Empirical-Exercise-1/main/social_and_task_network.csv"
data = fread(path, header = TRUE)
```

```{r}
# make adjacency matrix, subset to just the non-zero answers
nonzero_edges = data[social_tie > 0 | task_tie > 0]
full_network = graph.data.frame(nonzero_edges)
plot(full_network)
```
```{r}
#removing edges where there is no social tie to get social network
network_social = delete.edges(full_network, E(full_network)[get.edge.attribute(full_network,name = "social_tie")==0])
plot(network_social)
```
```{r}
#removing edges where there is no task tie to get task network
network_task = delete.edges(full_network, E(full_network)[get.edge.attribute(full_network,name = "task_tie")==0])
plot(network_task)
```
***
1. First, consider the social and task ties as two separate networks.
(A) Use igraph to generate indegree, outdegree, closeness, betweenness, and 
PageRank centrality statistics for each individual the social and task networks.
***

```{r, warning=FALSE}
#Social Network Calculations
degree(network_social, mode='in') #indegree

degree(network_social,mode='out') #outdegree

closeness(network_social) #closeness

betweenness(network_social) #betweenness

#removing those with no ties for centrality computations
cnetwork_social = delete.vertices(network_social, V(network_social)[degree(network_social)==0])
page_rank(cnetwork_social)#pagerank
```
```{r, warning=FALSE}
#Task Network Calculations
degree(network_task, mode='in') #indegree

degree(network_task,mode='out') #outdegree

closeness(network_task)#closeness

betweenness(network_task)#betweenness

#removing those with no ties for centrality computations
cnetwork_task = delete.vertices(network_task, V(network_task)[degree(network_task)==0])
page_rank(cnetwork_task)#page rank
```
***
(B) Compute the correlations of the five centrality measures you generate for the social
network with the five measures generated for the task network. Which measures in the
task network are most closely related to those in the socializing network? Name at least
one insight can you draw from the relationships between these five measures across the
two networks.
***
```{r warning=FALSE}
getNetStats=function(net)
{
  indegree = degree(net, mode = "in")
  outdegree = degree(net, mode='out')
  close = closeness(net)
  between = betweenness(net)
  page_rank = page_rank(net)
  name=V(net)$name
  stats=as.data.table(list(name = name, indegree = indegree, outdegree=outdegree, 
                           close = close, between = between, page_rank=page_rank$vector))
  return(stats)
}

social_table <- getNetStats(network_social)
task_table <- getNetStats(network_task)
colnames(social_table) <- paste('social', colnames(social_table), sep=' ')
colnames(task_table) <- paste('task', colnames(task_table), sep=' ')


cor(social_table[,-1],task_table[,-1])

```
***
1(B) Commentary: 
The highest correlated metric between both networks is betweenness. This suggests that people in this network who are effective social bridges also function as effective task bridges. This matches my lived experience of networks, where socializing and collaborating often intertwine. Another interesting result is the near-zero correlation between social page_rank and task closeness. This may, however, be because our network is not fully connected, resulting in issues with both the path of the 'random surfer' in the page rank algorithm and the measure of closeness.

***
***
2. Next, consider the social and task ties together, as two distinct types of ties comprising
one network. Suppose that a tie is strong if it is above the mean strength for that type,
conditional on the tie existing—that is, do not include weights of 0 in the calculation of the
mean.
(A) Under this definition, does the network satisfy Strong Triadic Closure? Come up with
a solution that illustrates this visually, in a plot.
***
```{r}
# mean social tie, zero-values excluded
smean <- mean(E(full_network)$social_tie[E(full_network)$social_tie!=0])

# mean task tie, zero-values excluded
tmean <- mean(E(full_network)$task_tie[E(full_network)$task_tie!=0])

#function from Stack Overflow to space nodes properly

plot(full_network, edge.color=ifelse(E(full_network)$social_tie>smean, rgb(1, 0, 0.2, 1), rgb(0,0.4,1,0.5)),  layout=layout_with_kk, edge.arrow.size = 0.5)
plot(full_network, edge.color=ifelse(E(full_network)$task_tie>tmean, rgb(.4, 0, 0.6, 1), rgb(0.3,0.3,0,0.5)),layout=layout_with_kk, edge.arrow.size = 0.5, add=TRUE)
```
***
2(A) Commentary:
In strong triadic closure, if there is a strong tie from one vertex to two other vertices, those 'other vertices' should also be connected. 
In the above plot, red represents strong social ties, and purple represents strong task ties. We can see vertex 19 has strong ties to 20 and 16, however, these nodes are not directly connected, which could the network violates strong triadic closure, however, in this case, 18 functions as a local bridge using weak ties to connect 20 and 16.
***
(B) Also come up with a solution that illustrates this programmatically, such as by giving
the number or proportion of ties that are in violation of Strong Triadic Closure.
***
```{r}
# creating df to check strong triadic closure
stc_df <- data.frame(nonzero_edges)
# creating strong and weak binary columns for each tie type
stc_df$soc_strong <- rep(0,length(stc_df$social_tie))
stc_df$soc_weak <- rep(0,length(stc_df$social_tie))
stc_df$task_strong <- rep(0,length(stc_df$task_tie))
stc_df$task_weak <- rep(0,length(stc_df$task_tie))

for (i in 1:length(stc_df$social_tie)){
  ifelse (stc_df$social_tie[i]>smean, stc_df$soc_strong[i] <- 1, stc_df$soc_weak[i] <- 1)
  ifelse (stc_df$task_tie[i]>tmean, stc_df$task_strong[i] <- 1, stc_df$task_weak[i] <- 1)
}
```
```{r}
# looping to check for strong closure
for (i in 1:22){
  strong_ego <- stc_df[which(stc_df$ego==i & (stc_df$soc_strong==1 | stc_df$task_strong==1)),]
  if (nrow(strong_ego)>1){
  ties <- as.vector(strong_ego$alter)
  perm_ties <- permut(ties,m=2)
    for (i in 1:nrow(perm_ties)) {
      vec <- as.vector(perm_ties[i,])
      index <- stc_df[which(stc_df$ego==vec[1] & stc_df$alter==vec[2]),]
      if(nrow(index)==0){print(c('ego',strong_ego$ego[1],'alter',vec[1],vec[2]))}
  }
  }
}
```
***
2(B). Commentary: 
Here we see the 12 ego->alter1 ego->alter2 groups where there is no alter1<->alter2 tie, thus violating strong triadic closure. These represent roughly 15% of our dataset of nonzero edges.
***
***
3. Continue to treat the social and task ties as two distinct types of ties comprising one network.
(A) It is also possible to compute betweenness on the edges in a network, as well as the
vertices. This is a good measure of the flow of information and resources through a
network. Calculate the edge-level betweenness for both the social and task ties.
***
```{r, warning=FALSE, message=FALSE, include=FALSE}
# social network betweenness
btw <- suppressMessages(
  cluster_edge_betweenness(network_social, E(network_social)$weight, directed=TRUE, edge.betweenness = TRUE,
                           modularity=FALSE))
btw_social_dt <- as.data.table(cbind(btw$removed.edges,btw$edge.betweenness))
colnames(btw_social_dt) <-  c('edge number', 'betweenness') 
```
```{r}
btw_social_dt
```
```{r, warning=FALSE, message=FALSE, include=FALSE}
# task network betweenness
btw2 <- suppressMessages(
  cluster_edge_betweenness(network_task, E(network_task)$weight, directed=TRUE, edge.betweenness = TRUE,
                           modularity=FALSE))
btw_task_dt <- as.data.table(cbind(btw2$removed.edges,btw2$edge.betweenness))
colnames(btw_task_dt) <-  c('edge number', 'betweenness') 
```
```{r}
setorder(btw_task_dt, -betweenness)
btw_task_dt
```
***
(B) Does it seem like edges with high betweenness tend to be strong or weak ties, according
the measure of tie strength we used in Question 2? Does this result make sense given
the shape of the networks?
***
```{r}
# social network betweeness and strength comparisons 
setorder(btw_social_dt, 'edge number')
social_edges <- as.data.table(cbind(btw_social_dt,attr(E(network_social),'vnames')))
colnames(social_edges) <- c('edge number', 'betweenness','vertices')
social_edges$ego <- as.integer(sub('\\|\\d*$','',social_edges$vertices))
social_edges$alter <- as.integer(sub('\\d*\\|', '', social_edges$vertices))
social_edges <- subset(social_edges, select=-vertices)
setorder(social_edges, 'ego')
social_edges_ties <- social_edges %>% left_join(stc_df)
setorder(social_edges_ties,-'betweenness')
social_edges_ties
```
```{r}
setorder(btw_task_dt, 'edge number')
task_edges <- as.data.table(cbind(btw_task_dt,attr(E(network_task),'vnames')))
colnames(task_edges) <- c('edge number', 'betweenness','vertices')
task_edges$ego <- as.integer(sub('\\|\\d*$','',task_edges$vertices))
task_edges$alter <- as.integer(sub('\\d*\\|', '', task_edges$vertices))
task_edges <- subset(task_edges, select=-vertices)
setorder(task_edges, 'ego')
task_edges_ties <- task_edges %>% left_join(stc_df)
setorder(task_edges_ties,-'betweenness')
task_edges_ties
```

***
3(B) Commentary: Looking at these networks, our most between edges had socially weak ties, but task strong ties. This appears to connect to the structure of our network, where strong social ties are largely focused on a very gregarious vertex, leaving most of the network bridged by weak ties, while strong task ties are more dispersed.
***
4. Continue to treat the social and task ties as two distinct types of ties comprising one network.
How many pairs of nodes do not have walks between one another? Find a solution that
performs this calculation directly on the matrix—it is possible to verify this solution via
igraph afterward using distances(). See also ?distances() for an explanation of igraph’s
related network distance functions.
***
```{r}

adj_matrix <- as_adjacency_matrix(full_network)
adj_matrix <- as.matrix(adj_matrix)
max_walk <-  adj_matrix %^% 56
no_walk_number <- 0
for (i in 1:21){
  no_walk_number <- no_walk_number+sum(max_walk[i,]==0)
}

paste(no_walk_number,'pairs have no walk between them')

```

***
5. The network-level measure of degree centrality is a good indicator of the dispersion of the
degree distribution in a network. Generate and plot a network in R in which the network-level
measure of degree centrality,

Cdegree =∑Cdegree(v^∗)−Cdegree(vi)/(n−1)(n−2)

is equal to 1, and another where it is equal to 0. Would this relationship hold true for these
networks for other measures of centrality, such as closeness or betweenness?
***